Stop Lying to Me: New Visual Tools to Choose the Most Honest Nonlinear Dimension Reduction

### 2025-06-26 motivationletter/motivationletter.qmd
motivation letter revised

Please consider our manuscript entitled *“Looking at Non-Linear Dimension Reductions as Models in the Data Space”* for publication in the *Journal of Computational and Graphical Statistics*.

@@

Our manuscript introduces an approach that treats NLDR outputs as models and visualizes them directly in the original high-dimensional space using tours (movies of linear projections). This allows analysts to assess how well the NLDR model fits across all subspaces, identify mismatches, and compare the summaries provided by different methods. The approach we describe offer a new, visually interpretable way to diagnose NLDR results and support more trustworthy use of these techniques in high-dimensional data analysis.

We believe that this work is a strong fit for *JCGS*, given the journal’s focus on innovative statistical graphics, computational methodology, and tools that help analysts better understand high-dimensional data. Our approach combines statistical modeling, dynamic visualization, and interactive diagnostics, contributing to the journal’s mission of advancing the use of computation and graphics in statistical practice.

>>>>

Please consider our manuscript entitled *“Stop Lying to Me: New Visual Tools to Choose the Most Honest Nonlinear Dimension Reduction”* for publication in the *Journal of Computational and Graphical Statistics*.

@@

Our manuscript introduces an approach that treats NLDR outputs as models and visualizes the models with the data in the original high-dimensional space using tours (movies of linear projections). Numerical diagnostics are also generated from the fitted model. These allow analysts to have support to help them choose an appropriate NLDR representation of their data.

We believe that this work is a strong fit for *JCGS*, given the journal’s focus on innovative statistical graphics, computational methodology, and tools that help analysts better understand high-dimensional data. Our approach combines statistical modeling, interactive and dynamic graphics, and numerical diagnostics, contributing to the journal’s mission of advancing the use of computation and graphics in statistical practice.
